---
title: "Extending Study 2 Results"
date: "2022-10-18"
output: html_notebook
---

```{r}
library(dplyr)
library(ggplot2)

theme_set(theme_classic())
```


Load clean data set

```{r}
load("../../data/clean/s2_data.RData")
```



```{r}
w12_subset <- df %>% filter(week == "week12")
n_w12 <- w12_subset %>% count(screen_id) %>% nrow()
n_nicotine <- w12_subset %>% filter(!is.na(carmine_nicotine)) %>% nrow()
```


`r n_w12` participants had CPD values at Week 12. Of those, `r n_nicotine` have usual brand nicotine values that map from Carmines et al. (2019).


```{r}
w12_nomiss <- w12_subset %>% 
  filter(!is.na(carmine_nicotine))
```




```{r}
w12_nomiss$carmine_nic_scaled <- scale(w12_nomiss$carmine_nicotine)
```




## Difference in Mean Week 12 CPD by Treatment Group

With original covariates: Baseline CPD, age, sex, menthol status, population

### Original Data set

```{r}
# study
ff_study <- study_mean ~ dose + baseline_cpd + screen_age + screen_sex + menthol_status + project
study_og_fit <- lm(ff_study, data = w12_subset)
summary(study_og_fit)
```


```{r}
# total
ff_total <- total_cpd ~ dose + baseline_cpd + screen_age + screen_sex + menthol_status + project
total_og_fit <- lm(ff_total, data = w12_subset)
summary(total_og_fit)
```


### Dataset Limited to only those with UB nicotine values


```{r}
# study
study_new_fit <- lm(ff_study, data = w12_nomiss)
summary(study_new_fit)
```



```{r}
# total CPD
total_new_fit <- lm(ff_total, data = w12_nomiss)
summary(total_new_fit)
```



## Difference in Mean Week 12 CPD with UB Nicotine

With original covariates: Baseline CPD, age, sex, menthol status, population

_plus_ UB nicotine

```{r}
ff_study_nic <- study_mean ~ dose + baseline_cpd + screen_age + screen_sex + menthol_status + project + carmine_nic_scaled

study_nic_fit <- lm(ff_study_nic, data = w12_nomiss)
summary(study_nic_fit)
```



```{r}
ff_total_nic <- total_cpd ~ dose + baseline_cpd + screen_age + screen_sex + menthol_status + project + carmine_nic_scaled

total_nic_fit <- lm(ff_total_nic, data = w12_nomiss)
summary(total_nic_fit)
```


```{r}
models <- list(
  "study_og" = study_og_fit,
  "study_new"=  study_new_fit,
  "study_nic" = study_nic_fit,
  "total_og" = total_og_fit,
  "total_new" = total_new_fit,
  "total_nic" = total_nic_fit
  )

get_coefs <- function(model_vector) {
  coefs <- data.frame()
  
  for (m in 1:length(model_vector)) {
     tmp <- as.data.frame(coef(model_vector[[m]]))
     tmp <- cbind(tmp, as.data.frame(coef(summary(model_vector[[m]]))[,4]) )
     tmp$model <- names(model_vector[m])
     tmp <- tmp %>% tibble::rownames_to_column()
     names(tmp) <- c("feature", "beta", "p_val", "model")
     coefs <- rbind(coefs, tmp)
  }
  
  coefs %>%
    mutate(sig = case_when(
      p_val < .001 ~ "***",
      p_val < .01 ~ "**",
      p_val < .05 ~ "*"
    ))
}

coefs <- get_coefs(models)

```



```{r}
coefs_plt <- coefs %>%
  filter(feature != "(Intercept)") %>%
  tidyr::separate(
    col = model,
    into = c("outcome", "model")
  ) %>%
  mutate(model = factor(model, levels = c("og", "new", "nic")))

coefs_plt %>%
  filter(outcome == "study") %>% 
  ggplot(aes(y = feature, x = beta, fill = model, label = sig)) +
  geom_col(position = "dodge") +
  geom_text(aes(x = beta + 0.5), position = position_dodge(width = 0.9)) +
  labs(
    title = "TCORS Study 2 GLMs"
  )

ggsave("../figures/full_glm.png", dpi = "retina")
  
```



## Correlation Between CPD and UB Nicotine Content

```{r}
w12_recode %>%
  mutate(log_nic = log(carmine_nicotine)) %>%
  select(carmine_nicotine, adherence) %>%
  plot()
  
  
#   
#   select(screen_id, trt_recode, carmine_nicotine, baseline_cpd, total_cpd, study_mean, delta_cpd) %>%
#   tidyr::pivot_longer(-c(screen_id, trt_recode, carmine_nicotine)) %>%
#   ggplot(aes(x = carmine_nicotine, y = value)) +
#   geom_point() +
#   ggpmisc::stat_poly_line(se = FALSE) +
#   ggpmisc::stat_poly_eq() +
#   facet_grid(rows = vars(name), cols = vars(trt_recode)) +
#   labs(
#     title = "Correlation* Between UB Nic. and CPD",
#     caption = "* Lack thereof",
#     y = "Avg. Cigarettes Per Day",
#     x = "Usual Brand Nicotine Content (mg/g)"
#   )
# ggsave("../figures/ub_nic_cor.png", dpi = "retina")

```








A bit of a hack-wrangle..


```{r}
# combine VLNC conditions
w12_recode <- df %>% 
  mutate(trt_recode = ifelse(dose == 0.4 | dose == 2.4, "treatment", "control")) %>%
  filter(week == "week12")
ff_total <- total_cpd ~ trt_recode + baseline_cpd + screen_age + screen_sex + menthol_status + project
anova_recode <- lm(ff_total, data = w12_recode)
car::Anova(anova_recode, type = "III")
summary(anova_recode)

write.csv(w12_recode, file = "../../data/clean/w12_recode.csv", row.names = FALSE)

```


```{r}
trt <- w12_recode[w12_recode$trt_recode == "treatment", ]
ctl <- w12_recode[w12_recode$trt_recode == "control", ]
x1 <- predict(anova_recode, trt)
x2 <- predict(anova_recode, ctl)

w12_recode$predicted <- predict(anova_recode, w12_recode)

effsize::cohen.d(w12_recode$predicted, w12_recode$trt_recode)

```



